\section{Water Models}\label{sec:water_models}

% Navier stokes, Different level of water depth, Spatial, Spectral domain,
% eulerian, lagrangian frameworks, talk about their visual accuracy, SLProject,
% 0 a.d.


\subsection{Navier Stokes Equations}\label{subsec:navier_stokes}
\subsection{Spatial Domain}\label{subsec:spatial_domain}
\subsection{Spectral Domain}\label{subsec:spectral_domain}
\subsection{Methods Used in Industry Productions}\label{subsec:methods_industry}

% FFT, Projected Grid, LOD,

\subsection{Candidate Applications for Embedding}\label{subsec:candidate_apps}

% SLProject, 0 a.d.

We now present two candidate applications which need a real-time water model:
SLProject and 0 A.D.

\subsubsection{SLProject}



\subsubsection{0 A.D.}

% Talk about the history of the project, what it is, how it is maintained,
% open source, game engine, the tickets and contacts with the devs.

0 A.D. is a real-time strategy game representing the 500 B.C to 500 A.D era. It
is developed by Wildfire Games, an independent game developpment studio. The
game is free, open source\footnote{liscenced under the GNU-GPL and CC BY-SA
licence} and runs on Windows, MacOS and Linux.

The developpment of the game began in 2000 when three modding groups whished to
create a free game engine. Up until 2009 the source code was accessible only to
members of Wildfire Games. They opened up the code in July 2009 due to the
fading interest of a mod which used 0 A.D.'s game engine and the lack of
programmers. From then on many contributions have been made, notabely one
redesigning the code base, to make new contributions easier. The project is
still actively maintained as of today, with the latesest alpha-release (number
22 codnamed ``Venustas'') dating from July 26,
2017~\autocite{wildfire0adproject,wildfire0adstory}.

% Code base and Engine


% Current water and open ticket

The current water implementation has been created by a student form the
University of Waterloo, Canada, in 2006 as a course project
~\autocite{zaharia2006cs}. Over the years it has been slightly modified and
optimized. 
